---
title: "homework_1"
author: "Efe Ahmet Guden"
date: "4/19/2021"
output: html_document
---

---
title: "Data Visualization with Economic Indicators"
author: "Efe Ahmet Guden"
date: "4/19/2021"
output: html_document
---
```{r setup,echo=FALSE,warning=FALSE,message=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(zoo)
library(EVDS)
library(ggplot2)
library(lubridate)
library(tidyverse)

devtools::install_github("algopoly/EVDS")

```

# Introduction
Recently, Turkish economy suffers from high inflation and high unemployment rates. Therefore, in this paper, three macroeconomic indicators of Turkish economy are inspected: TL per USD exchange rate, consumer price index and unemployment rate. The relationship between actual data and the search volume of related keywords are analyzed.
The time-series data of these indicators are imported from [EVDS](https://evds2.tcmb.gov.tr/). EVDS is the official data delivery webiste of Central Bank of the Republic of Turkey.
The time-series data of these indicator's search volumes are imported from [Google Trends](https://trends.google.com/trends/). The wikipedia definition of Google Trends is: "Google Trends is a public web facility of Google Inc., based on Google Search that shows how often a particular search-term is entered relative to the total search-volume across various regions of the world, and in various languages."
The scope of this analysis is from January 2018 to January 2021. The data from Google Trends are collected on a weekly basis. EVDS data for exchange rate are collected on a daily basis. EVDS data for consumer price index and unemployment rate are collected on a monthly basis.

# 1) Exchange Rate
Tl per USD exchange rate is an important measure for Turkish economy since most of the imported goods are purchased with USD. Thus, an increase in the exchange rate may cause an increase in consumer price index, which is an indicator of cost of living.
The relationship between actual exchange rate and the search volume of the keyword "dolar kuru" will be analyzed.

## 1.1) Importing Data from EVDS
[EVDS package](https://algopoly.github.io/EVDS/index.html.) is used for importing data from EVDS.
```{r,echo=FALSE}
set_evds_key("TX5OzOJpfZ")
dolar_import <- as.data.frame(get_series(series= c("TP.DK.USD.A.YTL"), start_date="15-01-2018" , end_date="15-01-2021"))
```
Exchange rate data on holidays and weekends are missing. Those rows have to be excluded from the data in order to perform an analysis.
```{r,echo=FALSE}
dolar_evds<- na.omit(dolar_import[,2:3])
```
```{r,echo=FALSE}
colnames(dolar_evds) <-  c("Date" , "ExchangeRate")
dolar_evds$Date <-as.Date(dolar_evds$Date, format= "%d-%m-%Y")
dolar_evds$ExchangeRate <- as.numeric(dolar_evds$ExchangeRate)
```

After some manipulations the data looks like this:
```{r}
head(dolar_evds)
```
## 1.2) Importing Data from Google Trends
The data from Google Trends are first downloaded as a csv file. Then the function read.csv() is used for importing the Google Trends data to R.

```{r, echo=FALSE, include=FALSE}
setwd("/Users/ahmetguden/Desktop")
dolar_trends <- as.data.frame(read_csv("dolarkuru.csv"))
```
```{r,echo=FALSE}
colnames(dolar_trends) <- c("Date" , "SearchVolume")
dolar_trends$Date <-  as.Date(dolar_trends$Date, format = "%m/%d/%y")
dolar_trends$SearchVolume <- as.numeric(dolar_trends$SearchVolume)
```
After some manipulations the data looks like this:
```{r}
head(dolar_trends)
```
## 1.3) Exchange Rate Data Visualization
The data are visualized with the ggplot() function, which is a function from the ggplot2() package.

This plot shows the TL per USD exchange rate over time.

```{r,echo=FALSE, warning=FALSE}
ggplot(dolar_evds, aes(Date,ExchangeRate)) +
  geom_line() + 
  labs(title="TL per USD over time" ,x="Date",y = "TL per USD") +
  scale_x_date(date_breaks = "2 months", limits = as.Date(c("2018-02-01", "2021-01-01"))) +
  theme(axis.text.x = element_text(angle = 90))
```

One can observe that there is an increasing trend. There is a dramatic increase in 2018 August.Then the exchange rate gradually decreased until 2019 April. Then there is a rise and fall in exchange rate levels between 2019 May and 2019 September. There is a rapid increase in 2020 March which lasts until 2020 June. The exchange rate level peaked at 2020 December with a dramatic increase.

The relationship between the actual exchange rate and the search volume for the keyword "dolar kuru" will be analyzed with the following plots.
```{r,echo=FALSE, fig.show="hold", out.width="50%", warning = F}
ggplot(dolar_evds, aes(Date,ExchangeRate)) +
  geom_line() + 
  labs(title="TL per USD over time" ,x="Date",y = "TL per USD") +
  scale_x_date(date_breaks = "2 months", limits = as.Date(c("2018-02-01", "2021-01-01"))) +
  theme(axis.text.x = element_text(angle = 90))
ggplot(dolar_trends, aes(Date,SearchVolume)) +
  geom_line() + 
  labs(title="Search Volume of 'Dolar Kuru' " ,x="Date",y = "Search Volume") +
  scale_x_date(date_breaks = "2 months",limits = as.Date(c("2018-02-01", "2021-01-01"))) +
  theme(axis.text.x = element_text(angle = 90))
```
The search volume of the keyword "dolar kuru" reached the maximum level in 2018 August with a high jump. This relates to the dramatic increase in the exchange rate in 2018 August. There is also a dramatic increase in the search volume data in 2020 December which is corresponding to the jump in the exchange rate in 2020 December.

# 2)Consumer Price Index

Consumer price index (CPI) is an indicator of cost of living. Inflation is calculated by the increase in CPI. Thus, one can state that there is high inflation, when the CPI increases rapidly.
The relationship between actual CPI and the search volume of the keyword "enflasyon" will be analyzed.

## 2.1) Importing Data from EVDS

[EVDS package](https://algopoly.github.io/EVDS/index.html.) is used for importing data from EVDS.
```{r,echo=FALSE}
set_evds_key("TX5OzOJpfZ")
cpi_import <- as.data.frame(get_series(series= c("TP.FG.J0"), start_date="15-01-2018" , end_date="15-01-2021"))
cpi <- cpi_import[,2:3]
titles_cpi <- c("Date" , "CPI")
colnames(cpi) <-  titles_cpi
char_dates  <- c("2018-01-15",  "2018-02-15" , "2018-03-15"  ,"2018-04-15"  ,"2018-05-15",  "2018-06-15" , "2018-07-15",  "2018-08-15"  ,"2018-09-15" ,
                 "2018-10-15", "2018-11-15", "2018-12-15" ,"2019-01-15" , "2019-02-15"  ,"2019-03-15"  ,"2019-04-15"  ,"2019-05-15"  ,"2019-06-15" ,
                 "2019-07-15" , "2019-08-15" , "2019-09-15",  "2019-10-15" ,"2019-11-15" ,"2019-12-15" ,"2020-01-15"  ,"2020-02-15"  ,"2020-03-15" ,
                 "2020-04-15"  ,"2020-05-15"  ,"2020-06-15"  ,"2020-07-15"  ,"2020-08-15"  ,"2020-09-15",  "2020-10-15" ,"2020-11-15" ,"2020-12-15",
                 "2021-01-15" )
dates <- as.Date(char_dates, format = "%Y-%m-%d")
cpi$Date <- dates
cpi$CPI  <- as.numeric(cpi$CPI)
```
After some manipulations the data looks like this:
```{r}
head(cpi)
```

## 2.2) Importing Data from Google Trends

The data from Google Trends are first downloaded as a csv file. Then the function read.csv() is used for importing the Google Trends data to R.

```{r, echo=FALSE, warning=FALSE,include=FALSE }
setwd("/Users/ahmetguden/Desktop")
enflasyon <- as.data.frame(read_csv("enflasyon.csv"))
```


```{r, echo=FALSE}
enflasyon$Date <- as.Date(enflasyon$Date, format = "%m/%d/%y")
enflasyon$SearchVolume <- as.numeric(enflasyon$SearchVolume)
```

After some manipulations the data looks like this: 
```{r}
head(enflasyon)
```

## 2.3) CPI Data Visualization

This plot shows the consumer price index over time.

```{r,echo=FALSE,warning=FALSE}
ggplot(cpi, aes(Date,CPI)) +
  geom_point() +
  geom_line() +
  labs(title = "Consumer Price Index over Time" , x="Date" ,y="CPI") +
  scale_x_date(date_breaks = "2 months",  limits = as.Date(c("2018-02-01", "2021-01-01"))) +
  theme(axis.text.x = element_text(angle = 90))
```

One can observe that there is an increasing trend. The increase in the consumer price index means that the price levels are increasing, thus the cost of living is increasing. Notice that there is a jump in 2018 September. Therefore, there is high inflation in 2018 September. The cause of this jump may be dramatic increase in the exchange rate which happened in 2018 August.

The relationship between the CPI and the search volume for the keyword "enflasyon" will be analyzed with the following plots.
```{r,echo=FALSE, fig.show="hold", out.width="50%", warning = FALSE}
ggplot(cpi, aes(Date,CPI)) +
  geom_point() +
  geom_line() +
  labs(title = "Consumer Price Index over Time" , x="Date" ,y="CPI") +
  scale_x_date(date_breaks = "2 months",  limits = as.Date(c("2018-02-01", "2021-01-01"))) +
  theme(axis.text.x = element_text(angle = 90))

ggplot(enflasyon, aes(Date,SearchVolume)) +
  geom_line() +
  labs(title="Search Volume of 'Enflasyon'" , x="Date" , y="Search Volume") +
  scale_x_date(date_breaks = "2 months", limits = as.Date(c("2018-02-01", "2021-01-01"))) +
  theme(axis.text.x = element_text(angle = 90))
```
The search volume for the keyword "enflasyon" tends to fluctuate. The search volume increases dramatically at the end of each quarter (months March, June, September, December). The search volume reached the maximum level at the end of 2018. This may be related to the jump in the CPI, which means high inflation, in 2018 September.

# 3) Unemployment Rate
Unemployment rate is an indicator which measures the ratio of number of unemployed people to the number of people who are in the labour force. Turkish economy suffers from high unemployment rates, which leads to problems both economically and socially.
The relationship between unemployment rate and the search volume for the keyword "işsizlik oranı" will be analyzed.

## 3.1) Importing Data from EVDS

[EVDS package](https://algopoly.github.io/EVDS/index.html.) is used for importing data from EVDS.
```{r,echo=FALSE}
set_evds_key("TX5OzOJpfZ")
unemp_import <- as.data.frame(get_series(series= c("TP.YISGUCU2.G8" ), start_date="15-01-2018" , end_date="15-01-2021"))
unemp_rate_evds <- unemp_import[,2:3]
```
```{r,echo=FALSE}
colnames(unemp_rate_evds) <- c("Date" , "UnemployementRate")
char_dates  <- c("2018-01-15",  "2018-02-15" , "2018-03-15"  ,"2018-04-15"  ,"2018-05-15",  "2018-06-15" , "2018-07-15",  "2018-08-15"  ,"2018-09-15" ,
           "2018-10-15", "2018-11-15", "2018-12-15" ,"2019-01-15" , "2019-02-15"  ,"2019-03-15"  ,"2019-04-15"  ,"2019-05-15"  ,"2019-06-15" ,
            "2019-07-15" , "2019-08-15" , "2019-09-15",  "2019-10-15" ,"2019-11-15" ,"2019-12-15" ,"2020-01-15"  ,"2020-02-15"  ,"2020-03-15" ,
           "2020-04-15"  ,"2020-05-15"  ,"2020-06-15"  ,"2020-07-15"  ,"2020-08-15"  ,"2020-09-15",  "2020-10-15" ,"2020-11-15" ,"2020-12-15",
            "2021-01-15" )
dates <- as.Date(char_dates, format = "%Y-%m-%d")
unemp_rate_evds$Date <- dates
unemp_rate_evds$UnemployementRate <- as.numeric(unemp_rate_evds$UnemployementRate)
```
After some manipulations the data looks like this: 
```{r}
head(unemp_rate_evds)
```

## 3.2) Importing Data from Google Trends

The data from Google Trends are first downloaded as a csv file. Then the function read.csv() is used for importing the Google Trends data to R.
```{r,echo=FALSE,include=FALSE}
setwd("/Users/ahmetguden/Desktop")
unemp_rate_trends <- as.data.frame(read_csv("unemprate.csv"))
```
```{r,echo=FALSE}
unemp_rate_trends$Date <- as.Date(unemp_rate_trends$Date,format = "%m/%d/%y")
unemp_rate_trends$SearchVolume <- as.numeric(unemp_rate_trends$SearchVolume)
```
After some manipulations the data looks like this: 
```{r}
head(unemp_rate_trends)
```

## 3.3) Unemployment Rate Data Visualization

This plot shows the unemployment rate over time.

```{r,echo=FALSE,warning=FALSE}
ggplot(unemp_rate_evds, aes(Date,UnemployementRate)) +
  geom_point() +
  geom_line() +
  labs(title = "Unemployment Rate over Time", x="Date" , y="Unemployment Rate") +
  scale_x_date(date_breaks = "2 months") +
  theme(axis.text.x = element_text(angle = 90))
```

One can observe that there is an increase in the unemployment rate between 2018 May and 2018 December. Afterwards, unemployment rate fluactuated around 13.5. There is a dramatic increase in 2018 December and in 2020 June.

The relationship between unemployment rate and the search volume for the keyword "işsizlik oranı" will be analyzed with the following plots.

```{r,echo=FALSE, fig.show="hold", out.width="50%", warning = FALSE}
ggplot(unemp_rate_evds, aes(Date,UnemployementRate)) +
  geom_point() +
  geom_line() +
  labs(title = "Unemployment Rate over Time", x="Date" , y="Unemployment Rate") +
  scale_x_date(date_breaks = "2 months") +
  theme(axis.text.x = element_text(angle = 90))

ggplot(unemp_rate_trends, aes(Date,SearchVolume)) +
  geom_line() +
  labs(title = "Search Volume of 'İşsizlik Oranı'" , x= "Date" , y= "Search Volume") +
  scale_x_date(date_breaks = "2 months", limits = as.Date(c("2018-02-01", "2021-01-01"))) +
  theme(axis.text.x = element_text(angle = 90))
```
There is no trend in the search volume of the keyword "işsizlik oranı". Search volume fluctuates. The maximum search volume is in 2020 June.This may be related to the jump in the unemployment rate in 2020 June. 

# Conclusion
Three macroeconomic indicators of Turkish economy is inspected. The relationship between the actual data and the search volume of related keywords are analyzed. One can observe that the dramatic increase in an economic indicator leads to a dramatic increase in search volume. To conclude, there is a relationship between actual data and the serach volume of related keywords.




